---
title: "Sociodemographic Drivers of Mental Health (Binary)"
author: "Emily Tyszka"
date: "`r Sys.Date()`"
linkcolor: blue
output:
  rmdformats::readthedown:
    highlight: kate
    number_sections: true
    code_folding: show
    df_print: paged
---

# Initial Setup and Package Loads in R 
```{r package loads, message = FALSE, comment = NA}

library(Epi); library(janitor); library(here); library(readxl); library(magrittr); 
library(haven); library(MASS); library(skimr); library(naniar); library(car); 
library(rms); library(broom); library(DescTools); library(psych); library(caret); 
library(GGally); library(knitr); library(tidyverse) 

```

# Load raw data. 

```{r load_data, warning = FALSE}

masterraw <- read_dta(file = "data.11132020.dta")

```


# Mutate variables to correct type
```{r mutate, warning = FALSE}

master<- masterraw %>% 
  mutate(USRace =as.factor(USRace)) %>%
  mutate(PatientSex = as.factor(PatientSex))%>%
  mutate(Householdincome = as.factor(Householdincome)) %>%
  mutate(LiveWithOtherPeople = as.factor(LiveWithOtherPeople)) %>%
  mutate(MSHealthCareProvider = as.factor(MSHealthCareProvider)) %>%
  mutate(MSHealthCareLocation = as.factor(MSHealthCareLocation)) %>%
  mutate(HighEducLevel = as.factor(HighEducLevel)) %>%
  mutate(MaritalStatus = as.factor(MaritalStatus)) %>%
  mutate(EmployStatus = as.factor(EmployStatus)) %>%
  mutate(BetterMentalHealth = as.factor(BetterMentalHealth)) 

```



# Summary of variables being investigated
|       Variable         |  Description                                                                       |
|------------------------|------------------------------------------------------------------------------------|
|   BetterMentalHealth   | OUTCOME - Mental health (0 = Poor/Fair Mental Health, 1 = Better Mental Health)    |
|        USRace          | Race (AI/NH, Asian, Black, Don't know, MiddleEastern, Multiracial, White)          | 
|      PatientSex        | Sex (F, M, Other)                                                                  |
|     USInsurance2       | Insurance Type (Public, None, Other, Private, Combination, VA)                       |
|  LiveWithOtherPeople   | Live alone or with others (LiveAlone, PreferNotAnswer, Yes )                       |
|  MSHealthCareLocation  | Provider of care (GenNeuroClinic, MSCenter, NoTreatmentLocation, OtherLocation, PrimaryCareCommunityHealth)|
|     HighEducLevel      | Highest level of education (AssociateOrTechnical, College, MastersPlus, LessThanHighSchool, HighSchool, SomeCollege)|
|     MaritalStatus      | Marital status (MarriedOrCohabitating, Unmarried, Separated)                       |
|     EmployStatus       | Employment status (Unemployed, Employed, Student)                                  |
|    Householdincome     | Yearly household income group (Less than $15,000, $15,000-$30,000, $30,001-$50,000, $50,001-$100,000, $100,001-$150,000, $150,001-$200,000, $200,001+) |
|   overallhealth_age    | Age at time of survey, in years                                                    |
|      AgefirstMS        | Age of first MS diagnosis, in years                                                |




# $\chi$^2^ and Fisher Tests

## Summary of $\chi$^2^ and Fisher Tests

|       Variable         |  p-value for $\chi$^2^ test  |  p-value for Fisher test  |
|------------------------|------------------------------|---------------------------|
| US_Race | 0.1373 | 0.1535 |
| PatientSex | 0.2273 | 0.2393 |
| USInsurance2 | **9.8e-11** | **0.0004998** |
| LiveWithOtherPeople | **0.004426** | **0.006497** |
| MSHealthCareLocation | **0.5862** | **0.5772** |
| HighEducLevel  | **2.118e-14** | **0.0004998** |
| MaritalStatus | **8.395e-05** | **0.0004998** |
| EmployStatus | **4.402e-13** | **0.0004998** |
| Householdincome | **7.201e-08** | **0.0004998** |


## US_Race
```{r race, warning = FALSE}
master <- master %>% 
  mutate(USRace = as.factor(USRace))

master$USRace <- fct_collapse(master$USRace,
  "AI/NH" = "AI/NH",
  Asian = "Asian",
  Black = "Black",
  DK = "DK",
  MiddleEastern = "MiddleEastern",
  Multiracial = "Multiracial",
  White = "White",
  NULL = "")

summary(master$USRace)

master %>%
  tabyl(BetterMentalHealth, USRace) %>%
  adorn_totals("row") %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting() %>%
  adorn_ns() %>%
  knitr::kable() 
```


```{r race tests, warning = FALSE}
chisq.test(master$BetterMentalHealth, master$USRace)

racetable <- table(master$BetterMentalHealth, master$USRace)
fisher.test(racetable)
```

## PatientSex
```{r sex, warning = FALSE}
master <- master %>% 
  mutate(PatientSex = as.factor(PatientSex))

master$PatientSex <- fct_collapse(master$PatientSex, NULL = c("", "PreferNotAnswer"))

summary(master$PatientSex)

master %>%
  tabyl(BetterMentalHealth, PatientSex) %>%
  adorn_totals("row") %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting() %>%
  adorn_ns() %>%
  knitr::kable() 
```


```{r sex tests, warning = FALSE}
chisq.test(master$BetterMentalHealth, master$PatientSex)

sextable <- table(master$BetterMentalHealth, master$PatientSex)
fisher.test(sextable)
```

## USInsurance2
```{r insurance levels, warning = FALSE}
master <- master %>% 
  mutate(USInsurance = as.factor(USInsurance))

summary(master$USInsurance)

master$USInsurance2 <- fct_collapse(master$USInsurance,
  Public = c("MEDICAID","MEDICARE","PUBLIC"),
  None = "NONE",
  Other = "OTHER",
  Private = "PRIVATE",
  VA = c("VA", "VA + PRIVATE","VA + PUBLIC + PRIVATE","VA + PUBLIC"),
  Combination = "PRIVATE + PUBLIC",
  NULL = c("DK", "PREFERNOTANSWER", NA, ""))
```

```{r insurance, warning = FALSE}

summary(master$USInsurance2)

master %>%
  tabyl(BetterMentalHealth, USInsurance2) %>%
  adorn_totals("row") %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting() %>%
  adorn_ns() %>%
  knitr::kable() 
```


```{r insurance tests, warning = FALSE}
chisq.test(master$BetterMentalHealth, master$USInsurance2)

insurtable <- table(master$BetterMentalHealth, master$USInsurance2)
fisher.test(insurtable, simulate.p.value = TRUE)
```

## LiveWithOtherPeople

```{r others, warning = FALSE}

master$LiveWithOtherPeople <- fct_collapse(master$LiveWithOtherPeople,
  LiveAlone = "LiveAlone",
  Yes = "Yes",
  PreferNotAnswer = "PreferNotAnswer",
  NULL = "")

summary(master$LiveWithOtherPeople)

master %>%
  tabyl(BetterMentalHealth, LiveWithOtherPeople) %>%
  adorn_totals("row") %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting() %>%
  adorn_ns() %>%
  knitr::kable() 
```

```{r others tests, warning = FALSE}
chisq.test(master$BetterMentalHealth, master$LiveWithOtherPeople)

otherstable <- table(master$BetterMentalHealth, master$LiveWithOtherPeople)
fisher.test(otherstable, simulate.p.value = TRUE)
```


## MSHealthcareLocation

```{r location, warning = FALSE}

master$MSHealthCareLocation <- fct_collapse(master$MSHealthCareLocation,
  NULL = "")

summary(master$MSHealthCareLocation)

master %>%
  tabyl(BetterMentalHealth, MSHealthCareLocation) %>%
  adorn_totals("row") %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting() %>%
  adorn_ns() %>%
  knitr::kable() 
```

```{r location tests, warning = FALSE}
chisq.test(master$BetterMentalHealth, master$MSHealthCareLocation)

locationtable <- table(master$BetterMentalHealth, master$MSHealthCareLocation)
fisher.test(locationtable, simulate.p.value = TRUE)
```


## HighEducLevel

```{r education, warning = FALSE}

master$HighEducLevel <- fct_collapse(master$HighEducLevel,
  LessThanHighSchool = "ElementaryMiddleSchool",
  HighSchool = "HighSchoolGraduate",
  AssociateOrTechnical = c("AssociateDegree", "TechnicalDegree"),
  SomeCollege = "SomeCollege",
  College = "BachelorsDegree",
  MastersPlus = c("DoctoralDegree", "MastersDegree"),
  NULL = c("NA","DontKnow" ))

summary(master$HighEducLevel)

master %>%
  tabyl(BetterMentalHealth, HighEducLevel) %>%
  adorn_totals("row") %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting() %>%
  adorn_ns() %>%
  knitr::kable() 
```

```{r education tests, warning = FALSE}
chisq.test(master$BetterMentalHealth, master$HighEducLevel)

edutable <- table(master$BetterMentalHealth, master$HighEducLevel)
fisher.test(edutable, simulate.p.value = TRUE)
```

## MaritalStatus

```{r marital, warning = FALSE}

master$MaritalStatus <- fct_collapse(master$MaritalStatus,
  Unmarried = c("NeverMarried", "Divorced", "Widowed"),
  MarriedOrCohabitating = c("Married", "CohabitationPartner"),
  NULL = c("NA","PreferNotAnswer", ""))

summary(master$MaritalStatus)

master %>%
  tabyl(BetterMentalHealth, MaritalStatus) %>%
  adorn_totals("row") %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting() %>%
  adorn_ns() %>%
  knitr::kable() 
```

```{r marital tests, warning = FALSE}
chisq.test(master$BetterMentalHealth, master$MaritalStatus)

maritaltable <- table(master$BetterMentalHealth, master$MaritalStatus)
fisher.test(maritaltable, simulate.p.value = TRUE)
```


## EmployStatus

```{r employment, warning = FALSE}

master$EmployStatus <- fct_collapse(master$EmployStatus,
  Unemployed = c("Unemployed", "UnemployedNotLooking", "Disabled", "WorkmensComp", "Homemaker", "Retired"),
  Employed = c("Employed", "EmployedOutside", "EmployedHome"),
  AssociateOrTechnical = c("AssociateDegree", "TechnicalDegree"),
  NULL = c("NA","PreferNotAnswer", ""))

summary(master$EmployStatus)

master %>%
  tabyl(BetterMentalHealth, EmployStatus) %>%
  adorn_totals("row") %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting() %>%
  adorn_ns() %>%
  knitr::kable() 
```

```{r employment tests, warning = FALSE}
chisq.test(master$BetterMentalHealth, master$EmployStatus)

employtable <- table(master$BetterMentalHealth, master$EmployStatus)
fisher.test(employtable, simulate.p.value = TRUE)
```

## Householdincome 

```{r income, warning = FALSE}

master$Householdincome <- fct_collapse(master$Householdincome, 
                                       "Less than $15,000" = "1",
                                       "$15,000-$30,000" = "2",
                                       "$30,001-$50,000" = "3",
                                       "$50,001-$100,000" = "4",
                                       "$100,001-$150,000" = "5",
                                       "$150,001-$200,000" = "6",
                                       "$200,001+" = "7",
                                       NULL = "8", "")

summary(master$Householdincome)

master %>%
  tabyl(BetterMentalHealth, Householdincome) %>%
  adorn_totals("row") %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting() %>%
  adorn_ns() %>%
  knitr::kable() 
```

```{r income tests, warning = FALSE}
chisq.test(master$BetterMentalHealth, master$Householdincome)

incometable <- table(master$BetterMentalHealth, master$Householdincome)
fisher.test(incometable, simulate.p.value = TRUE)
```


# Ordinal Logistic Regression Models

## Summaries of the AgeFirstMS and OVERALLHEALTH_AGE Variables
```{r summary age first MS, warning = FALSE}

summary(master$AgefirstMS)
ggplot(data=master, aes(x = AgefirstMS))+
  geom_histogram()

```

```{r summary age, warning = FALSE}

summary(master$OVERALLHEALTH_AGE)
ggplot(data=master, aes(x = OVERALLHEALTH_AGE))+
  geom_histogram()
 
```


## Generalized Linear Model for BetterMentalHealth Variable

```{r regression model ages 1, warning = FALSE}

ages1 <- glm(BetterMentalHealth ~ AgefirstMS + OVERALLHEALTH_AGE, data = master, family = "binomial")
summary(ages1)


```



