title: "MS Healthcare Behavior Analysis"
author: "Emily Tyszka"
date: "`r Sys.Date()`"
linkcolor: blue
output:
  rmdformats::readthedown:
    highlight: kate
    number_sections: true
    code_folding: show
    df_print: paged
---


# Initial Setup and Package Loads in R 
```{r package loads, message = FALSE, comment = NA}

library(Epi); library(janitor); library(here); library(readxl);
 library(skimr); library(naniar); library(car); library(rms); library(broom); 
library(caret); library(GGally); library(knitr); library(tidyverse) 

```

# Load raw data. 

```{r load_data, warning = FALSE}
masterraw <- read_excel("master.112020.usa.ms.hc.xlsx") %>% 
  tbl_df %>% clean_names()
```

# Mutate variables to correct type
```{r mutate, warning = FALSE}
master1 <- masterraw %>% 
  mutate(pdds = as.factor(pdds)) %>% 
  mutate(m_sbestcharacterizes = as.factor(m_sbestcharacterizes)) %>% 
  mutate(race =as.factor(race)) %>%
  mutate(patient_sex = as.factor(patient_sex))%>%
  mutate(ethnicity = as.factor(ethnicity)) %>%
  mutate(ms_health_care_facility = as.factor(ms_health_care_facility)) %>%
  mutate(ms_health_care_location = as.factor(ms_health_care_location))%>%
  mutate(us_race = as.factor(us_race))
  
```

# Collapse Down Variables
## Race
```{r see levels, warning = FALSE}
summary(master1$race)
```

```{r collapse races multi, warning = FALSE}
master1$race2 <- fct_collapse(master1$race,
  AfricanAmerican = "AfricanAmerican",
  AmericanIndian = "AmericanIndian",
  MiddleEastern = "MiddleEastern",
  Asian = c("SouthAsian",'SouthAsian, OtherAsian', 'OtherAsian'),
  White = "White",
  DontKnow = "DontKnowEthnicity",
  TwoOrMoreRaces = c('AmericanIndian, AfricanAmerican, DontKnowEthnicity',
                        'AmericanIndian, MiddleEastern, White','AmericanIndian, SouthAsian',
                        'AmericanIndian, White',
                        'NativeHawaiian, White',
                        'OtherAsian, NativeHawaiian, White, DontKnowEthnicity',
                        'AfricanAmerican, White', 
                        'AmericanIndian, AfricanAmerican',
                        'AmericanIndian, AfricanAmerican, White',
                        'AmericanIndian, OtherAsian, AfricanAmerican, White',
                        'AmericanIndian, SouthAsian, White',
                        'OtherAsian, AfricanAmerican, White',
                        'MiddleEastern, White',
                        'OtherAsian, White' ))
master1$race2 <- factor(master1$race2, ordered=TRUE, levels =c("AfricanAmerican", "AmericanIndian","Asian","MiddleEastern","White","DontKnow","TwoOrMoreRaces"))
summary(master1$race2)
```
## Ethnicity
```{r collapse ethnicity , warning = FALSE}
master1$ethnicity <- fct_collapse(master1$ethnicity,
  Hispanic = "Hispanic",
  NonHispanic = "NonHispanic",
  NULL = c("DontKnow"))
```

## Sex 
```{r collapse sex , warning = FALSE}
master1$patient_sex <- fct_collapse(master1$patient_sex,
  Male = "M",
  Female = "F",
  NULL = c("Other","PreferNotAnswer"))
```

## Facility 
```{r collapse facilities, warning = FALSE}
master1$ms_health_care_facility2 <- fct_collapse(master1$ms_health_care_facility,
  CancerCenter = c("CANCER CENTER","Cancer Center at hospital"),
  Doctor= c("Doctor's office", "Doctors office","dr office", 
            "dr. office", "Dr Office or Hospital for Solumedrol"),
  Home = "Home",
  Hospital = c("Hospital","Mayo Clinic", "University of Penn", "University"),
  InfusionCenter = c("Infusion Center", "Infusion Center", "infusion suite"),
  InternalMedicine = "Internal medicine",
  Kaiser = "Kaiser",
  NeurologyPractice = c("Neuro Office","Neuro's practice", "Neurologist office", 
                        "neurologist within my health care system", "neurologist's office",
                        "Neurology practice", "OFFICE OF NEUROLOGIST", 
                        "MS Center of Washington - solo practice of MS neurologist"),
  Opthamologist = "Opthamologist",
  Physiatrist = "physiatrist practice",
  PCP = c("PRIMARY CARE PHYSICIAN", "private dr. office or PCP", "private practice", 
          "Private practice", "private practice office"),
  VAHospital= c("VA Hospital MS Center","VA Hospital", "VAMC"),
  NULL = "don't understand question")


summary(master1$ms_health_care_facility2)
```

# Race

```{r race summary , warning = FALSE}
summary(master1$race2)
```

## Provider
```{r provider race plot, warning = FALSE}

ggplot(data=master1, aes(x = usms_health_care_provider, fill = usms_health_care_provider))+
  geom_bar()+
  facet_wrap(~race2)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master1, aes(x = race2, fill = race2))+
  geom_bar()+
  facet_wrap(~usms_health_care_provider)+ theme(axis.text.x = element_text(angle = 90))
```


```{r provider race, warning = FALSE}
master1 %>% 
    tabyl(race2, usms_health_care_provider) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master1$race2, master1$usms_health_care_provider)
```
## Facility

### With NA's

```{r facility race plot, warning = FALSE}

ggplot(data=master1, aes(x = ms_health_care_facility2, fill = ms_health_care_facility2))+
  geom_bar()+
  facet_wrap(~race2)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master1, aes(x = race2, fill = race2))+
  geom_bar()+
  facet_wrap(~ms_health_care_facility2)+ theme(axis.text.x = element_text(angle = 90))
```

```{r race facility, warning = FALSE}
master1 %>% 
    tabyl(race2, ms_health_care_facility2) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master1$race2, master1$ms_health_care_facility2)
```

### Without NA's

```{r facilty race plot no na, warning = FALSE}
master5<- drop_na(master1, ms_health_care_facility2)

ggplot(data=master5, aes(x = ms_health_care_facility2, fill = ms_health_care_facility2))+
  geom_bar()+
  facet_wrap(~race2)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master5, aes(x = race2, fill = race2))+
  geom_bar()+
  facet_wrap(~ms_health_care_facility2)+ theme(axis.text.x = element_text(angle = 90))
```

```{r race facility no na, warning = FALSE}
master5 %>% 
    tabyl(race2, ms_health_care_facility2) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master5$race2, master5$ms_health_care_facility2)
```

## Location
```{r location race plot, warning = FALSE}

ggplot(data=master1, aes(x = usms_health_care_location, fill = usms_health_care_location))+
  geom_bar()+
  facet_wrap(~race2)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master1, aes(x = race2, fill = race2))+
  geom_bar()+
  facet_wrap(~usms_health_care_location)+ theme(axis.text.x = element_text(angle = 90))
```


```{r location race, warning = FALSE}
master1 %>% 
    tabyl(race2, usms_health_care_location) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master1$race2, master1$usms_health_care_location)
```



# US Race

```{r us race summary , warning = FALSE}
summary(master1$us_race)
```

## Provider
```{r provider us race plot, warning = FALSE}

ggplot(data=master1, aes(x = usms_health_care_provider, fill = usms_health_care_provider))+
  geom_bar()+
  facet_wrap(~us_race)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master1, aes(x = us_race, fill = us_race))+
  geom_bar()+
  facet_wrap(~usms_health_care_provider)+ theme(axis.text.x = element_text(angle = 90))
```


```{r provider us race, warning = FALSE}
master1 %>% 
    tabyl(us_race, usms_health_care_provider) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master1$us_race, master1$usms_health_care_provider)
```
## Facility

### With NA's

```{r facility us race plot, warning = FALSE}

ggplot(data=master1, aes(x = ms_health_care_facility2, fill = ms_health_care_facility2))+
  geom_bar()+
  facet_wrap(~us_race)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master1, aes(x = us_race, fill = us_race))+
  geom_bar()+
  facet_wrap(~ms_health_care_facility2)+ theme(axis.text.x = element_text(angle = 90))
```

```{r us race facility, warning = FALSE}
master1 %>% 
    tabyl(us_race, ms_health_care_facility2) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master1$us_race, master1$ms_health_care_facility2)
```

### Without NA's

```{r facility us race plot no na, warning = FALSE}
master5<- drop_na(master1, ms_health_care_facility2)

ggplot(data=master5, aes(x = ms_health_care_facility2, fill = ms_health_care_facility2))+
  geom_bar()+
  facet_wrap(~us_race)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master5, aes(x = us_race, fill = us_race))+
  geom_bar()+
  facet_wrap(~ms_health_care_facility2)+ theme(axis.text.x = element_text(angle = 90))
```

```{r us race facility no na, warning = FALSE}
master5 %>% 
    tabyl(us_race, ms_health_care_facility2) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master5$us_race, master5$ms_health_care_facility2)
```

## Location
```{r location us race plot, warning = FALSE}

ggplot(data=master1, aes(x = usms_health_care_location, fill = usms_health_care_location))+
  geom_bar()+
  facet_wrap(~us_race)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master1, aes(x = us_race, fill = us_race))+
  geom_bar()+
  facet_wrap(~usms_health_care_location)+ theme(axis.text.x = element_text(angle = 90))
```


```{r location us race, warning = FALSE}
master1 %>% 
    tabyl(us_race, usms_health_care_location) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master1$us_race, master1$usms_health_care_location)
```


# Disability

```{r disability summary , warning = FALSE}
summary(master1$pdds)
```

## Provider

```{r provider disability plot, warning = FALSE}
master2 <-drop_na(master1, pdds)

ggplot(data=master2, aes(x = usms_health_care_provider, fill = usms_health_care_provider))+
  geom_bar()+
  facet_wrap(~pdds)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master2, aes(x = pdds, fill = pdds))+
  geom_bar()+
  facet_wrap(~usms_health_care_provider)+ theme(axis.text.x = element_text(angle = 90))
```

```{r provider disability, warning = FALSE}

master2 %>% 
    tabyl(usms_health_care_provider, pdds) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master2$usms_health_care_provider, master2$pdds)
```

## Facility
### With NA's
```{r facility disability plot, warning = FALSE}

ggplot(data=master1, aes(x = ms_health_care_facility2, fill = ms_health_care_facility2))+
  geom_bar()+
  facet_wrap(~pdds)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master1, aes(x =pdds, fill = pdds))+
  geom_bar()+
  facet_wrap(~ms_health_care_facility2)+ theme(axis.text.x = element_text(angle = 90))
```

```{r disability facility, warning = FALSE}
master1 %>% 
    tabyl(pdds, ms_health_care_facility2) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master1$pdds, master1$ms_health_care_facility2)
```

### Without NA's
```{r facility disability plot no na, warning = FALSE}

ggplot(data=master5, aes(x = ms_health_care_facility2, fill = ms_health_care_facility2))+
  geom_bar()+
  facet_wrap(~pdds)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master5, aes(x =pdds, fill = pdds))+
  geom_bar()+
  facet_wrap(~ms_health_care_facility2)+ theme(axis.text.x = element_text(angle = 90))
```

```{r disability facility no na, warning = FALSE}
master5 %>% 
    tabyl(pdds, ms_health_care_facility2) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master5$pdds, master5$ms_health_care_facility2)
```

## Location

```{r location disability plot, warning = FALSE}

ggplot(data=master2, aes(x = usms_health_care_location, fill = usms_health_care_location))+
  geom_bar()+
  facet_wrap(~pdds)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master2, aes(x = pdds, fill = pdds))+
  geom_bar()+
  facet_wrap(~usms_health_care_location)+ theme(axis.text.x = element_text(angle = 90))
```

```{r location disability, warning = FALSE}

master2 %>% 
    tabyl(usms_health_care_location, pdds) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master2$usms_health_care_location, master2$pdds)
```

# MS Type

```{r type summary , warning = FALSE}
summary(master1$m_sbestcharacterizes)
```

## Provider

```{r provider type plot , warning = FALSE}
ggplot(data=master1, aes(x = usms_health_care_provider, fill = usms_health_care_provider))+
  geom_bar()+
  facet_wrap(~m_sbestcharacterizes)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master1, aes(x = m_sbestcharacterizes, fill = m_sbestcharacterizes))+
  geom_bar()+
  facet_wrap(~usms_health_care_provider)+ theme(axis.text.x = element_text(angle = 90))
```

```{r provider type, warning = FALSE}

master1 %>% 
    tabyl(usms_health_care_provider, m_sbestcharacterizes) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master1$usms_health_care_provider, master1$m_sbestcharacterizes)

```

## Facility
### With NA's
```{r facility type plot, warning = FALSE}

ggplot(data=master1, aes(x = ms_health_care_facility2, fill = ms_health_care_facility2))+
  geom_bar()+
  facet_wrap(~m_sbestcharacterizes)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master1, aes(x =m_sbestcharacterizes, fill = m_sbestcharacterizes))+
  geom_bar()+
  facet_wrap(~ms_health_care_facility2)+ theme(axis.text.x = element_text(angle = 90))
```

```{r type facility, warning = FALSE}
master1 %>% 
    tabyl(m_sbestcharacterizes, ms_health_care_facility2) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master1$m_sbestcharacterizes, master1$ms_health_care_facility2)
```

### Without NA's
```{r facility type plot no na, warning = FALSE}

ggplot(data=master5, aes(x = ms_health_care_facility2, fill = ms_health_care_facility2))+
  geom_bar()+
  facet_wrap(~m_sbestcharacterizes)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master5, aes(x =m_sbestcharacterizes, fill = m_sbestcharacterizes))+
  geom_bar()+
  facet_wrap(~ms_health_care_facility2)+ theme(axis.text.x = element_text(angle = 90))
```

```{r type facility no na, warning = FALSE}
master5 %>% 
    tabyl(m_sbestcharacterizes, ms_health_care_facility2) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master5$m_sbestcharacterizes, master5$ms_health_care_facility2)
```

## Location

```{r location type plot , warning = FALSE}
ggplot(data=master1, aes(x = usms_health_care_location, fill = usms_health_care_location))+
  geom_bar()+
  facet_wrap(~m_sbestcharacterizes)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master1, aes(x = m_sbestcharacterizes, fill = m_sbestcharacterizes))+
  geom_bar()+
  facet_wrap(~usms_health_care_location)+ theme(axis.text.x = element_text(angle = 90))
```

```{r location type, warning = FALSE}

master1 %>% 
    tabyl(usms_health_care_location, m_sbestcharacterizes) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master1$usms_health_care_location, master1$m_sbestcharacterizes)

```

# Sex

```{r sex summary , warning = FALSE}
summary(master1$patient_sex)
  
master3 <- drop_na(master1, patient_sex)
summary(master3$patient_sex)
```

## Provider

```{r provider sex plot , warning = FALSE}
ggplot(data=master3, aes(x = usms_health_care_provider, fill = usms_health_care_provider))+
  geom_bar()+
  facet_wrap(~patient_sex)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master3, aes(x = patient_sex, fill = patient_sex))+
  geom_bar()+
  facet_wrap(~usms_health_care_provider)+ theme(axis.text.x = element_text(angle = 90))
```

```{r provider sex, warning = FALSE}

master3 %>% 
    tabyl(usms_health_care_provider, patient_sex) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master3$usms_health_care_provider, master3$patient_sex)
```

## Facility
### With NA's
```{r facility sex plot, warning = FALSE}

ggplot(data=master3, aes(x = ms_health_care_facility2, fill = ms_health_care_facility2))+
  geom_bar()+
  facet_wrap(~patient_sex)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master3, aes(x =patient_sex, fill = patient_sex))+
  geom_bar()+
  facet_wrap(~ms_health_care_facility2)+ theme(axis.text.x = element_text(angle = 90))
```

```{r sex facility, warning = FALSE}
master3 %>% 
    tabyl(patient_sex, ms_health_care_facility2) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master3$patient_sex, master3$ms_health_care_facility2)
```

### Without NA's
```{r facility sex plot no na, warning = FALSE}

master5 <- drop_na(master3, ms_health_care_facility2)

ggplot(data=master5, aes(x = ms_health_care_facility2, fill = ms_health_care_facility2))+
  geom_bar()+
  facet_wrap(~patient_sex)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master5, aes(x =patient_sex, fill = patient_sex))+
  geom_bar()+
  facet_wrap(~ms_health_care_facility2)+ theme(axis.text.x = element_text(angle = 90))
```

```{r sex facility no na, warning = FALSE}
master5 %>% 
    tabyl(patient_sex, ms_health_care_facility2) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master5$patient_sex, master5$ms_health_care_facility2)
```
## Location

```{r location sex plot , warning = FALSE}
ggplot(data=master3, aes(x = usms_health_care_location, fill = usms_health_care_location))+
  geom_bar()+
  facet_wrap(~patient_sex)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master3, aes(x = patient_sex, fill = patient_sex))+
  geom_bar()+
  facet_wrap(~usms_health_care_location)+ theme(axis.text.x = element_text(angle = 90))
```

```{r location sex, warning = FALSE}

master3 %>% 
    tabyl(usms_health_care_location, patient_sex) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master3$usms_health_care_location, master3$patient_sex)
```

# Hispanic / Non-Hispanic

```{r hisp summary , warning = FALSE}

summary(master1$ethnicity)

master4 <- drop_na(master1, ethnicity)
summary(master4$ethnicity)

```

## Provider

```{r provider hisp plot , warning = FALSE}
ggplot(data=master4, aes(x = usms_health_care_provider, fill = usms_health_care_provider))+
  geom_bar()+
  facet_wrap(~ethnicity)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master4, aes(x = ethnicity, fill = ethnicity))+
  geom_bar()+
  facet_wrap(~usms_health_care_provider)+ theme(axis.text.x = element_text(angle = 90))
```

```{r provider hisp, warning = FALSE}

master4 %>% 
    tabyl(usms_health_care_provider, ethnicity) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master4$usms_health_care_provider, master4$ethnicity)

```

## Facility
### With NA's
```{r facility hisp plot, warning = FALSE}

ggplot(data=master4, aes(x = ms_health_care_facility2, fill = ms_health_care_facility2))+
  geom_bar()+
  facet_wrap(~ethnicity)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master4, aes(x =ethnicity, fill = ethnicity))+
  geom_bar()+
  facet_wrap(~ms_health_care_facility2)+ theme(axis.text.x = element_text(angle = 90))
```

```{r hisp facility, warning = FALSE}
master4 %>% 
    tabyl(ethnicity, ms_health_care_facility2) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master4$ethnicity, master4$ms_health_care_facility2)
```

### Without NA's
```{r facility hisp plot no na, warning = FALSE}

ggplot(data=master5, aes(x = ms_health_care_facility2, fill = ms_health_care_facility2))+
  geom_bar()+
  facet_wrap(~ethnicity)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master5, aes(x =ethnicity, fill = ethnicity))+
  geom_bar()+
  facet_wrap(~ms_health_care_facility2)+ theme(axis.text.x = element_text(angle = 90))
```

```{r hisp facility no na, warning = FALSE}
master5 %>% 
    tabyl(ethnicity, ms_health_care_facility2) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master5$ethnicity, master5$ms_health_care_facility2)
```

## Location

```{r location hisp plot , warning = FALSE}
ggplot(data=master4, aes(x = usms_health_care_location, fill = usms_health_care_location))+
  geom_bar()+
  facet_wrap(~ethnicity)+ theme(axis.text.x = element_text(angle = 90))

ggplot(data=master4, aes(x = ethnicity, fill = ethnicity))+
  geom_bar()+
  facet_wrap(~usms_health_care_location)+ theme(axis.text.x = element_text(angle = 90))
```

```{r location hisp, warning = FALSE}

master4 %>% 
    tabyl(usms_health_care_location, ethnicity) %>%
    adorn_totals(where = c("row", "col")) %>% # add row, column totals
    kable 

chisq.test(master4$usms_health_care_location, master4$ethnicity)

```





# Summary of $\chi$^2^ Tests

|                    | Race    | US Race | Disability |    Sex   |   MS Type  | Hispanic |
|--------------------|---------|---------|------------|----------|------------|----------|
| Facility           | 0.8796  |  0.8796 |  0.7464    |  0.2024  |   0.5672   |  0.08885 |
| Provider           | 0.9101  |  0.9457 |  0.06418   |  0.4811  | **0.0391** |  0.5267  |
| Location           | 0.9545  |  0.9445 |  0.2046    |  0.51    |**0.01765** |  0.9825  |

# Session Information

```{r}
sessioninfo::session_info()
```
